---
title: "109_cse"
output: html_document
date: "2026-03-03"
---

### installing packages

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
```

### importing data

```{r}
raw_pilot_data <- read_csv("C:/Users/haran/metalab/109_cse/cse_109_pilot_data.csv")
```

### data cleaning

#### flagging practice trials

```{r}
raw_pilot_data = raw_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(block_marker = coalesce(str_detect(stimulus, regex("Blokk\\s*1\\s*kezdődik", ignore_case = TRUE)),
                                 FALSE),
         start_seen = cummax(trial_index == 0),
         marker_seen = cummax(block_marker),
         trial_exp = if_else(start_seen == 1 & marker_seen == 0, "practice", "experimental")) %>% 
  ungroup()
```

#### flagging previous trials

```{r}
raw_pilot_data = raw_pilot_data %>%
  mutate(prev_congruency = lag(congruency, n = 4),
         prev_correct = lag(correct, n = 4),
         correct = ifelse(correct == "TRUE", 1, 0),
         prev_correct = ifelse(prev_correct == "TRUE", 1, 0),
         first_trial = ifelse(lag(str_detect(stimulus, "Blokk \\d+ kezdődik"), 4) == TRUE, 1, 0),
         rt = as.numeric(rt))
```

#### filtering subjects with accuracy below \<60%

```{r}
accuracy = raw_pilot_data %>% 
  filter(
    task == "probe") %>%
  filter(
    trial_exp != "practice",
    first_trial != 1) %>%
      group_by(subj_code) %>%
      summarize(all_accuracy = mean(correct, na.rm = TRUE))

raw_pilot_data = raw_pilot_data %>% 
  left_join(accuracy, by = "subj_code") %>%
  filter(all_accuracy > 0.60)
```

#### filtering data

```{r}
filtered_pilot_data = raw_pilot_data %>%
  filter(
    trial_exp != "practice",
    task == "probe",
    first_trial != 1,
    correct == 1,
    prev_correct == 1,
    rt >= 150 | is.na(rt))
```

```{r}
processed_pilot_data = filtered_pilot_data %>% 
  select(
    -stimulus,
    -response,
    -trial_type,
    -trial_index,
    -response_index,
    -plugin_version,
    -time_elapsed,
    -success,
    -question_order,
    -correct_response,
    -block_marker,
    -marker_seen,
    -prev_correct,
    -correct,
    -start_seen,
    -trial_exp,
    -first_trial,
    -name)
```

### extracting demographic data

```{r}
demography_data = raw_pilot_data %>%
  filter(trial_type %in% c("survey-text", "survey-multi-choice"))
    
demography_data = demography_data %>%
  select(response, subj_code)

demography_data = demography_data %>%
  mutate(age = str_extract(response, '"age":"\\d+"') %>%
      str_extract("\\d+") %>%
      as.numeric(),
    
    gender = str_extract(response, '"gender":"[^"]+"') %>%
      str_remove_all('"gender":"') %>%
      str_remove('"'))

demography_data = demography_data %>%
  select(-response)

demography_data = demography_data %>%
  group_by(subj_code) %>%
  summarize(
    age = coalesce(age[1], age[2]),
    gender = coalesce(gender[1], gender[2]))
```
