---
title: "109_cse"
output: html_document
date: "2026-03-03"
---

### Installing packages

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
```

### Importing data

```{r}
raw_pilot_data = read_csv("raw_pilot_data.csv")
```

### Data cleaning

### Extracting demographic data

```{r}
demography_data = raw_pilot_data %>%
  filter(trial_type %in% c("survey-text", "survey-multi-choice"))
    
demography_data = demography_data %>%
  select(response, subj_code)

demography_data = demography_data %>%
  mutate(age = str_extract(response, '"age":"\\d+"') %>%
      str_extract("\\d+") %>%
      as.numeric(),
    
    gender = str_extract(response, '"gender":"[^"]+"') %>%
      str_remove_all('"gender":"') %>%
      str_remove('"'))

demography_data = demography_data %>%
  select(-response)

demography_data = demography_data %>%
  group_by(subj_code) %>%
  summarize(
    age = coalesce(age[1], age[2]),
    gender = coalesce(gender[1], gender[2]))
```

#### Flagging practice trials

```{r}
raw_pilot_data = raw_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(block_marker = coalesce(str_detect(stimulus, regex("Blokk\\s*1\\s*kezdődik", ignore_case = TRUE)),
                                 FALSE),
         start_seen = cummax(trial_index == 0),
         marker_seen = cummax(block_marker),
         trial_exp = if_else(start_seen == 1 & marker_seen == 0, "practice", "experimental")) %>% 
  ungroup()
```

#### Flagging previous trials

```{r}
raw_pilot_data = raw_pilot_data %>%
  mutate(prev_congruency = lag(congruency, n = 4),
         prev_correct = lag(correct, n = 4),
         correct = ifelse(correct == "TRUE", 1, 0),
         prev_correct = ifelse(prev_correct == "TRUE", 1, 0),
         first_trial = ifelse(lag(str_detect(stimulus, "Blokk \\d+ kezdődik"), 4) == TRUE, 1, 0),
         rt = as.numeric(rt))
```

### Flagging diagnostic trials

```{r}
raw_pilot_data = raw_pilot_data %>% 
  filter(task=="probe")%>% 
  filter(trial_exp == "experimental")

raw_pilot_data = raw_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(trial_role = if_else(monetary == 1, "inducer", "diagnostic")) %>% 
  ungroup()
```

### Flagging monetary-monetary sequence (the trials were generated by a randomization python code with specific parameters. It is possible, that because of this randomness, it happened in a few cases that a monetary trial followed another monetary trial. This shouldn't be the case, therefore we exclude these pairs)

```{r}
raw_pilot_data = raw_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(double_monetary = if_else(
  monetary == 1 & (lag(monetary, default = 0) == 1 | lead(monetary, default = 0) == 1),
  1, 0)) %>% 
  filter(double_monetary != 1) %>% 
  ungroup()
```

```{r}
raw_pilot_data %>% 
  count(double_monetary)
```

### Filtering subjects with accuracy below 60%

```{r}
accuracy = raw_pilot_data %>% 
      group_by(subj_code) %>%
      summarize(all_accuracy = mean(correct, na.rm = TRUE))

raw_pilot_data = raw_pilot_data %>% 
  select(-any_of("all_accuracy")) %>% 
  left_join(accuracy, by = "subj_code") %>%
  filter(all_accuracy > 0.60, first_trial != 1) 
```

```{r}
#raw_pilot_data %>% 
  #count(color)
```

#### Filtering data

```{r}
filtered_pilot_data = raw_pilot_data %>%
  filter(correct == 1,
         prev_correct == 1,
         rt >= 150 | is.na(rt))
```

```{r}
#filtered_pilot_data %>% 
  #count(color)
```


```{r}
processed_pilot_data = filtered_pilot_data %>% 
  select(
    -double_monetary,
    -stimulus,
    -response,
    -trial_type,
    -trial_index,
    -plugin_version,
    -time_elapsed,
    -success,
    -question_order,
    -correct_response,
    -block_marker,
    -marker_seen,
    -prev_correct,
    -start_seen,
    -trial_exp,
    -first_trial,
    -name)
```

### Flagging previous color & labels
```{r}
processed_pilot_data = processed_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(prev_color = lag(color)) %>% 
  ungroup()
```

```{r}
processed_pilot_data = processed_pilot_data %>% 
  mutate(prev_color = case_when(
    prev_color == "red" ~ "loss",
    prev_color == "green" ~ "gain",
    prev_color %in% c("magenta", "blue", "yellow") ~ "neutral")) %>% 
  filter(!is.na(prev_color))
```

### Descriptive analysis function

```{r}
dstats = function(x, na.omit=FALSE){
  if (na.omit)
  x = x[!is.na(x)]
  min = min(x)
  max = max(x)
  m = mean(x)
  median = median(x)
  n = length(x)
  stds = sd(x)
  skew = 3*(m-median)/stds
  kurt = (length(x) * sum((x - mean(x))^4)) / (sum((x - mean(x))^2)^2) - 3

  
#Return all calculated values to vectors
  return(c(
    min = min,
    max = max,
    mean = m,
    median = median,
    sd = stds,
    skewness = skew,
    kurtosis = kurt
  ))
}
```

### Descriptive statistics for aggregated RT values

```{r}
mean_rt_df = processed_pilot_data %>% 
  group_by(subj_code) %>% 
  summarise(mean_rt_congruent = mean(rt[congruency == "congruent"]),
            mean_rt_incongruent = mean(rt[congruency == "incongruent"]))

myvars = c("mean_rt_congruent", "mean_rt_incongruent")
sapply(mean_rt_df[myvars], dstats)
```

### Density plot

```{r}
density_rt = ggplot(processed_pilot_data, aes(rt))
density_rt + geom_density()
```

### Violinplots for CE

#### Appropriate data structure for visualization

```{r}
ce_rt = processed_pilot_data %>% 
  group_by(subj_code, congruency) %>% 
  summarise(mean_rt = mean(rt))
```

```{r}
violin = ggplot(ce_rt, aes(congruency, mean_rt, fill = congruency)) +
  geom_violin(trim = TRUE, bounds = c(0, Inf), width = 0.5) +
  geom_boxplot(width = 0.1, colour = c("black", "white"), outlier.shape = NA) +
  labs(title = "Congruency Effect", x = "Congruency", y = "Mean reaction time (ms)") +
  papaja::theme_apa(base_family = "Times") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("congruent" = "#FAC775", "incongruent" = "#BA7517"))
#ggsave("ce_violin.png", plot = violin, width = 16, height = 9, units = "cm", dpi = 320)
print(violin)
```

### Participant level CSE

#### Appropriate data structure for visualizing the CSE

```{r}
cse_df = processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency) %>% 
  summarise(mean_rt = mean(rt))

cse_descriptive_df = cse_df %>% 
  group_by(congruency, prev_congruency) %>% 
  summarise(N = n(),
            mean = mean(mean_rt),
            sd = sd(mean_rt),
            se = sd/sqrt(N))
```

#### Participant CSE

```{r}
wide_df = cse_df %>% 
  pivot_wider(names_from = c(prev_congruency, congruency),
              values_from = mean_rt)

subject_cse = wide_df %>% 
  mutate(CSE = (congruent_incongruent - congruent_congruent) - (incongruent_incongruent - incongruent_congruent))

ggplot(subject_cse, aes(CSE)) +
  geom_density(fill = "#011526", alpha = 0.25) +
  scale_x_continuous(breaks = seq(floor(min(subject_cse$CSE)),
                                  ceiling(max(subject_cse$CSE)),
                                  by = 5)) +
  theme_minimal()
```

```{r}
ggplot(subject_cse, aes(CSE)) +
  geom_histogram(fill = "#011526", alpha = 0.25, binwidth = 10) +
  scale_x_continuous(breaks = seq(floor(min(subject_cse$CSE)),
                                  ceiling(max(subject_cse$CSE)),
                                  by = 5)) +
  theme_minimal()
```

### Plot for CSE

```{r}
cse_plot = ggplot(cse_descriptive_df, aes(prev_congruency, mean, colour = congruency, group = congruency)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1.2) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 1.2) +
                      labs("Congruency-sequence Effect",
                       x = "Previous Congruency",
                       y = "Mean reaction time (ms)",
                       colour = "Congruency") + 
  scale_colour_manual(values = c("#FAC775", "#BA7517")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right",
                        plot.title = element_text(hjust = 0.5, face = "bold"),
                        panel.grid = element_blank(),
                        axis.line = element_line(color = "grey", linewidth = 1))
#ggsave("overall_cse_plot.png", plot = cse_plot, width = 16, height = 9, units = "cm", dpi = 320)
print(cse_plot)
```

### CSE descriptive

```{r}
myvars = c("CSE")
sapply(subject_cse[myvars], dstats)
```

### Participant level CSE for the three different condition

```{r}
cse_color_df= processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency, prev_color) %>% 
  summarise(mean_rt = mean(rt))

wide_color_df = cse_color_df %>% 
  pivot_wider(names_from = c(prev_color, prev_congruency, congruency),
              values_from = mean_rt)

subject_color_cse = wide_color_df %>% 
  mutate(CSE_loss = (loss_congruent_incongruent - loss_congruent_congruent) - (loss_incongruent_incongruent - loss_incongruent_congruent),
         CSE_gain = (gain_congruent_incongruent - gain_congruent_congruent) - (gain_incongruent_incongruent - gain_incongruent_congruent),
         CSE_neutral = (neutral_congruent_incongruent - neutral_congruent_congruent) - (neutral_incongruent_incongruent - neutral_incongruent_congruent)) %>% 
  filter(!is.na(CSE_loss) & !is.na(CSE_gain) & !is.na(CSE_neutral))
```

### Descriptive stats for all CSE conditions

```{r}
myvars2 = c("CSE_loss", "CSE_gain", "CSE_neutral")
sapply(subject_color_cse[myvars2], dstats)
```

### Visualizing the CSE distributions

```{r}
subject_color_cse %>%
  pivot_longer(cols = c(CSE_loss, CSE_gain, CSE_neutral),
               names_to = "condition",
               values_to = "CSE") %>%
  filter(!is.na(CSE)) %>%
  ggplot(aes(x = CSE, fill = condition)) +
  geom_histogram(alpha = 0.25, binwidth = 5, fill = "#011526") +
  facet_wrap(~condition) +
  scale_x_continuous(breaks = seq(-200, 200, by = 25))
  theme_minimal()
```

### Plotting CSE with emotional condition

```{r}
cse_color_summary_df = processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency, prev_color) %>% 
  summarise(mean_rt = mean(rt))

cse_color_plot_df = cse_color_summary_df %>% 
  group_by(congruency, prev_congruency, prev_color) %>% 
  summarise(N = n(),
            grand_mean_rt = mean(mean_rt),
            sd_rt = sd(mean_rt),
            se_rt = sd_rt/sqrt(N))
```

```{r}
cse_color_plot_df %>% 
  ggplot()+
  aes(y = grand_mean_rt, x = prev_congruency, colour = prev_color, group = congruency)+
  geom_point(size = 3) +
  geom_path(linewidth = 1.2)+
  geom_errorbar(aes(ymin = grand_mean_rt - se_rt, ymax = grand_mean_rt + se_rt), width = 0.2, linewidth = 0.8) +
                  labs(title = "Congruency-sequence Effect",
                       x = "Previous Congruency",
                       y = "Mean reaction time (ms)",
                       colour = "Congruency") + 
                  scale_colour_manual(values = c("#03AC13", "#B90E0A", "#616569")) +
                  theme_minimal(base_size = 14) +
                  theme(legend.position = "right",
                        plot.title = element_text(hjust = 0.5, face = "bold"),
                        panel.grid = element_blank(),
                        axis.line = element_line(color = "grey", linewidth = 1)) +
  facet_grid(cols = vars (prev_color))
#ggsave("cse_color_plot.png", plot = last_plot(), width = 16, height = 9, units = "cm", dpi = 320)
```

## Analysis

### Normality for CSE rt

```{r}
shapiro.test(subject_color_cse$CSE_loss)
shapiro.test(subject_color_cse$CSE_gain)
shapiro.test(subject_color_cse$CSE_neutral)
```

### Assumption check for H1

```{r}
assumption_h1 = with(mean_rt_df, mean_rt_incongruent - mean_rt_congruent)
shapiro.test(assumption_h1)
```

#### If the assumption of normality is not met, we will use Wilcoxon signed-rank test for H1

```{r}
wilcox.test(mean_rt_df$mean_rt_incongruent, mean_rt_df$mean_rt_congruent, paired = TRUE, alternative = "greater")
```

### H1 - Congruency Effect (paired samples t-test)

```{r}
t.test(mean_rt_df$mean_rt_incongruent, mean_rt_df$mean_rt_congruent, paired = T, alternative = "greater")
```

### H2 - Congruency-sequence Effect (one samples t-test)

```{r}
t.test(subject_cse$CSE, mu = 0, alternative = "greater")
```

### Assumption check for H3

```{r}
assumption_h3 = with(subject_color_cse, CSE_loss - CSE_neutral)
shapiro.test(assumption_h3)
```

#### If the assumption of normality is not met, we will use Wilcoxon signed-rank test for H2

```{r}
wilcox.test(subject_color_cse$CSE_loss, subject_color_cse$CSE_neutral, paired = T, alternative = "greater")
```


### H3 - CSE will be stronger in the loss condition compared to neutral (paired samples t-test)

```{r}
t.test(subject_color_cse$CSE_loss, subject_color_cse$CSE_neutral, paired = T, alternative = "greater")
```

### Assumption check for H4

```{r}
assumption_h4 = with(subject_color_cse, CSE_neutral - CSE_gain)
shapiro.test(assumption_h4)
```

#### If the assumption of normality is not met, we will use Wilcoxon signed-rank test for H2

```{r}
wilcox.test(subject_color_cse$CSE_gain, subject_color_cse$CSE_neutral, paired = T, alternative = "less")
```

### H4 - CSE will be weaker in the gain condition compared to neutral (paired samples t-test)

```{r}
t.test(subject_color_cse$CSE_gain, subject_color_cse$CSE_neutral, paired = T, alternative = "less")
```



## Bayesian inference using Dienes's bayes factor function

### Store the H1 test parameters in an object

```{r}
t_test_H1 = t.test(mean_rt_df$mean_rt_incongruent, mean_rt_df$mean_rt_congruent, paired = T, alternative = "greater")
```

```{r}
t_test_H1
source("dienes_bayes.R")
Bf_h1 = old_Bf(sd = t_test_H1$stderr,
   obtained = t_test_H1$estimate,
   dfdata = t_test_H1$parameter,
   meanoftheory = 0,
   sdtheory = 63,
   dftheory = 10^10,
   tail = 1)
print(Bf_h1)

h1_range = seq(from=3,to=1000,by=10) 
range_test = t(sapply(h1_range, function(sdi) {
  bf_val = old_Bf(
    sd = t_test_H1$stderr,
    obtained = t_test_H1$estimate,
    dfdata = t_test_H1$parameter,
    meanoftheory = 0,
    sdtheory = sdi,
    dftheory = 10^10,
    tail = 1)
  c(sdtheory = sdi, BF = bf_val)
  })) 

colnames(range_test) = c ("sdtheory", "BF")

ev_for_h1 <- subset(data.frame(range_test), BF > 3)
low_threshold_mcm <- min(ev_for_h1$sdtheory)
high_threshold_mcm <- max(ev_for_h1$sdtheory)
print(range_test)
```

## Repeat the previous steps for all focal hypothesis

### H2

```{r}
t_test_H2 = t.test(subject_cse$CSE, mu = 0, alternative = "greater")
```

```{r}
t_test_H2
Bf_h2 = old_Bf(sd = t_test_H2$stderr,
   obtained = t_test_H2$estimate,
   dfdata = t_test_H2$parameter,
   meanoftheory = 0,
   sdtheory = 30,
   dftheory = 10^10,
   tail = 1)
print(Bf_h2)

h2_range = seq(from=3,to=1000,by=10) 
range_test = t(sapply(h2_range, function(sdi) {
  bf_val = old_Bf(
    sd = t_test_H2$stderr,
    obtained = t_test_H2$estimate,
    dfdata = t_test_H2$parameter,
    meanoftheory = 0,
    sdtheory = sdi,
    dftheory = 10^10,
    tail = 1)
  c(sdtheory = sdi, BF = bf_val)
  })) 

colnames(range_test) = c ("sdtheory", "BF")

ev_for_h2 <- subset(data.frame(range_test), BF > 3)
low_threshold_mcm <- min(ev_for_h2$sdtheory)
high_threshold_mcm <- max(ev_for_h2$sdtheory)
print(range_test)
```

### H3

```{r}
t_test_H3 = t.test(subject_color_cse$CSE_loss, subject_color_cse$CSE_neutral, paired = T, alternative = "greater")
```

```{r}
BF_h3 = old_Bf(sd = t_test_H3$stderr,
   obtained = t_test_H3$estimate, 
   dfdata = t_test_H3$parameter,
   meanoftheory = 0,
   sdtheory = 15,  
   dftheory = 10^10, 
   tail = 1)
BF_h3

h3_range = seq(from=3,to=50,by=1) 
range_test = t(sapply(h3_range, function(sdi) {
  bf_val = old_Bf(
    sd = t_test_H3$stderr,
    obtained = t_test_H3$estimate,
    dfdata = t_test_H3$parameter,
    meanoftheory = 0,
    sdtheory = sdi,
    dftheory = 10^10,
    tail = 1)
  c(sdtheory = sdi, BF = bf_val)
  })) 

colnames(range_test) = c ("sdtheory", "BF")

ev_for_h3 <- subset(data.frame(range_test), BF > 3)
low_threshold_mcm <- min(ev_for_h3$sdtheory)
high_threshold_mcm <- max(ev_for_h3$sdtheory)
print(range_test)
```

### H4 

```{r}
t_test_H4 = t.test(subject_color_cse$CSE_gain, subject_color_cse$CSE_neutral, paired = T, alternative = "less")
```

```{r}
BF_h4 = old_Bf(sd = t_test_H4$stderr,
   obtained = t_test_H4$estimate, 
   dfdata = t_test_H4$parameter,
   meanoftheory = 0,
   sdtheory = 15,  
   dftheory = 10^10, 
   tail = 1)
BF_h4

h4_range = seq(from=3,to=50,by=1) 
range_test = t(sapply(h4_range, function(sdi) {
  bf_val = old_Bf(
    sd = t_test_H4$stderr,
    obtained = t_test_H4$estimate,
    dfdata = t_test_H4$parameter,
    meanoftheory = 0,
    sdtheory = sdi,
    dftheory = 10^10,
    tail = 1)
  c(sdtheory = sdi, BF = bf_val)
  })) 

colnames(range_test) = c ("sdtheory", "BF")

ev_for_h4 <- subset(data.frame(range_test), BF > 3)
low_threshold_mcm <- min(ev_for_h4$sdtheory)
high_threshold_mcm <- max(ev_for_h4$sdtheory)
print(range_test)
```




