---
title: "109_cse"
output: html_document
date: "2026-03-03"
---

### installing packages

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
```

### importing data

```{r}
raw_pilot_data = read_csv("inlab_pilot_109_source.csv")
```

### data cleaning

#### flagging practice trials

```{r}
raw_pilot_data = raw_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(block_marker = coalesce(str_detect(stimulus, regex("Blokk\\s*1\\s*kezdődik", ignore_case = TRUE)),
                                 FALSE),
         start_seen = cummax(trial_index == 0),
         marker_seen = cummax(block_marker),
         trial_exp = if_else(start_seen == 1 & marker_seen == 0, "practice", "experimental")) %>% 
  ungroup()
```

#### flagging previous trials

```{r}
raw_pilot_data = raw_pilot_data %>%
  mutate(prev_congruency = lag(congruency, n = 4),
         prev_correct = lag(correct, n = 4),
         correct = ifelse(correct == "TRUE", 1, 0),
         prev_correct = ifelse(prev_correct == "TRUE", 1, 0),
         first_trial = ifelse(lag(str_detect(stimulus, "Blokk \\d+ kezdődik"), 4) == TRUE, 1, 0),
         rt = as.numeric(rt))
```

#### filtering subjects with accuracy below \<60%

```{r}
accuracy = raw_pilot_data %>% 
  filter(task == "probe") %>%
  filter(trial_exp != "practice", first_trial != 1) %>%
      group_by(subj_code) %>%
      summarize(all_accuracy = mean(correct, na.rm = TRUE))

raw_pilot_data = raw_pilot_data %>% 
  left_join(accuracy, by = "subj_code") %>%
  filter(all_accuracy > 0.60)


```

#### filtering data

```{r}
filtered_pilot_data = raw_pilot_data %>%
  filter(trial_exp != "practice",
         task == "probe",
         correct == 1,
         prev_correct == 1,
         rt >= 150 | is.na(rt))
```

```{r}
processed_pilot_data = filtered_pilot_data %>% 
  select(
    -stimulus,
    -response,
    -trial_type,
    -trial_index,
    -response_index,
    -plugin_version,
    -time_elapsed,
    -success,
    -question_order,
    -correct_response,
    -block_marker,
    -marker_seen,
    -prev_correct,
    -correct,
    -start_seen,
    -trial_exp,
    -first_trial,
    -name)
```

### Flagging previous color & labels
```{r}
processed_pilot_data = processed_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(prev_color = lag(color)) %>% 
  ungroup()
```

```{r}
processed_pilot_data = processed_pilot_data %>% 
  mutate(prev_color = case_when(
    prev_color == "red" ~ "loss",
    prev_color == "green" ~ "gain",
    prev_color %in% c("magenta", "blue", "yellow") ~ "neutral")) %>% 
  filter(!is.na(prev_color))
```

```{r}
processed_pilot_data %>% 
  
```

### extracting demographic data

```{r}
demography_data = raw_pilot_data %>%
  filter(trial_type %in% c("survey-text", "survey-multi-choice"))
    
demography_data = demography_data %>%
  select(response, subj_code)

demography_data = demography_data %>%
  mutate(age = str_extract(response, '"age":"\\d+"') %>%
      str_extract("\\d+") %>%
      as.numeric(),
    
    gender = str_extract(response, '"gender":"[^"]+"') %>%
      str_remove_all('"gender":"') %>%
      str_remove('"'))

demography_data = demography_data %>%
  select(-response)

demography_data = demography_data %>%
  group_by(subj_code) %>%
  summarize(
    age = coalesce(age[1], age[2]),
    gender = coalesce(gender[1], gender[2]))
```

### descriptive analysis function

```{r}
dstats = function(x, na.omit=FALSE){
  if (na.omit)
  x = x[!is.na(x)]
  min = min(x)
  max = max(x)
  m = mean(x)
  median = median(x)
  n = length(x)
  stds = sd(x)
  skew = 3*(m-median)/stds
  kurt = (length(x) * sum((x - mean(x))^4)) / (sum((x - mean(x))^2)^2) - 3

  
#Return all calculated values to vectors
  return(c(
    min = min,
    max = max,
    mean = m,
    median = median,
    sd = stds,
    skewness = skew,
    kurtosis = kurt
  ))
}
```

### Descriptive statistics for aggregated RT values

```{r}
mean_rt_df = processed_pilot_data %>% 
  group_by(subj_code) %>% 
  summarise(mean_rt_congruent = mean(rt[congruency == "congruent"]),
            mean_rt_incongruent = mean(rt[congruency == "incongruent"]))

myvars = c("mean_rt_congruent", "mean_rt_incongruent")
sapply(mean_rt_df[myvars], dstats)
```

### density plot

```{r}
density_rt = ggplot(processed_pilot_data, aes(rt))
density_rt + geom_density()
```

### Violinplots for CE

#### Appropriate data structure for visualization

```{r}
ce_rt = processed_pilot_data %>% 
  group_by(subj_code, congruency) %>% 
  summarise(mean_rt = mean(rt))
```

```{r}
violin = ggplot(ce_rt, aes(congruency, mean_rt, fill = congruency)) +
  geom_violin(trim = TRUE, bounds = c(0, Inf), width = 0.5) +
  geom_boxplot(width = 0.1, colour = c("black", "white"), outlier.shape = NA) +
  labs(title = "Congruency Effect", x = "Congruency", y = "Mean reaction time (ms)") +
  papaja::theme_apa(base_family = "Times") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("congruent" = "#FAC775", "incongruent" = "#BA7517"))
#ggsave("ce_violin.png", plot = violin, width = 16, height = 9, units = "cm", dpi = 320)
print(violin)
```

### Participant level CSE

#### Appropriate data structure for visualizing the CSE

```{r}
cse_df = processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency) %>% 
  summarise(mean_rt = mean(rt))

cse_descriptive_df = cse_df %>% 
  group_by(congruency, prev_congruency) %>% 
  summarise(N = n(),
            mean = mean(mean_rt),
            sd = sd(mean_rt),
            se = sd/sqrt(N))
```

#### Participant CSE

```{r}
wide_df = cse_df %>% 
  pivot_wider(names_from = c(prev_congruency, congruency),
              values_from = mean_rt)

subject_cse = wide_df %>% 
  mutate(CSE = (congruent_incongruent - congruent_congruent) - (incongruent_incongruent - incongruent_congruent))

ggplot(subject_cse, aes(CSE)) +
  geom_density(fill = "#011526", alpha = 0.25) +
  scale_x_continuous(breaks = seq(floor(min(subject_cse$CSE)),
                                  ceiling(max(subject_cse$CSE)),
                                  by = 5)) +
  theme_minimal()
```

```{r}
ggplot(subject_cse, aes(CSE)) +
  geom_histogram(fill = "#011526", alpha = 0.25, binwidth = 10) +
  scale_x_continuous(breaks = seq(floor(min(subject_cse$CSE)),
                                  ceiling(max(subject_cse$CSE)),
                                  by = 5)) +
  theme_minimal()
```

### Plot for CSE

```{r}
cse_plot = ggplot(cse_descriptive_df, aes(prev_congruency, mean, colour = congruency, group = congruency)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1.2) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 1.2) +
                      labs("Congruency-sequence Effect",
                       x = "Previous Congruency",
                       y = "Mean reaction time (ms)",
                       colour = "Congruency") + 
  scale_colour_manual(values = c("#FAC775", "#BA7517")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right",
                        plot.title = element_text(hjust = 0.5, face = "bold"),
                        panel.grid = element_blank(),
                        axis.line = element_line(color = "grey", linewidth = 1))
#ggsave("overall_cse_plot.png", plot = cse_plot, width = 16, height = 9, units = "cm", dpi = 320)
print(cse_plot)
```

### CSE descriptive

```{r}
myvars = c("CSE")
sapply(subject_cse[myvars], dstats)
```

### Participant level CSE for the three different condition

```{r}
cse_color_df= processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency, prev_color) %>% 
  summarise(N = n(),
            mean_rt = mean(rt))

wide_color_df = cse_color_df %>% 
  pivot_wider(names_from = c(prev_color, prev_congruency, congruency),
              values_from = mean_rt)

subject_color_cse = wide_color_df %>% 
  mutate(CSE_loss = (loss_congruent_incongruent - loss_congruent_congruent) - (loss_incongruent_incongruent - loss_incongruent_congruent),
         CSE_gain = (gain_congruent_incongruent - gain_congruent_congruent) - (gain_incongruent_incongruent - gain_incongruent_congruent),
         CSE_neutral = (neutral_congruent_incongruent - neutral_congruent_congruent) - (neutral_incongruent_incongruent - neutral_incongruent_congruent))

```

### Plotting CSE with emotional condition

```{r}
cse_color_summary_df = processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency, prev_color) %>% 
  summarise(mean_rt = mean(rt))

cse_color_plot_df = cse_color_summary_df %>% 
  group_by(congruency, prev_congruency, prev_color) %>% 
  summarise(N = n(),
            grand_mean_rt = mean(mean_rt),
            sd_rt = sd(mean_rt),
            se_rt = sd_rt/sqrt(N))
```

```{r}
cse_color_plot_df %>% 
  ggplot()+
  aes(y = grand_mean_rt, x = prev_congruency, colour = prev_color, group = congruency)+
  geom_point(size = 3) +
  geom_path(linewidth = 1.2)+
  geom_errorbar(aes(ymin = grand_mean_rt - se_rt, ymax = grand_mean_rt + se_rt), width = 0.2, linewidth = 0.8) +
                  labs(title = "Congruency-sequence Effect",
                       x = "Previous Congruency",
                       y = "Mean reaction time (ms)",
                       colour = "Congruency") + 
                  scale_colour_manual(values = c("#03AC13", "#B90E0A", "#E0E0E0")) +
                  theme_minimal(base_size = 14) +
                  theme(legend.position = "right",
                        plot.title = element_text(hjust = 0.5, face = "bold"),
                        panel.grid = element_blank(),
                        axis.line = element_line(color = "grey", linewidth = 1)) +
  facet_grid(cols = vars (prev_color))
#ggsave("cse_color_plot.png", plot = last_plot(), width = 16, height = 9, units = "cm", dpi = 320)
```

## Analysis

### Normality for CSE rt

```{r}
shapiro.test(subject_color_cse$CSE_loss)
shapiro.test(subject_color_cse$CSE_gain)
shapiro.test(subject_color_cse$CSE_neutral)
```

### Assumption check for H1

```{r}
assumption_h1 = with(mean_rt_df, mean_rt_incongruent - mean_rt_congruent)
shapiro.test(assumption_h1)
```

### H1 - Congruency Effect (paired samples t-test)

```{r}
t.test(mean_rt_df$mean_rt_incongruent, mean_rt_df$mean_rt_congruent, paired = T, alternative = "greater")
```

### H2 - Congruency-sequence Effect (one samples t-test)

```{r}
t.test(subject_cse$CSE, mu = 0, alternative = "greater")
```

### H2.1 - Congruency-sequence Effect for gain condition

```{r}
t.test(subject_color_cse$CSE_gain, mu = 0, alternative = "greater")
```

### H2.2 - Congruency-sequence Effect for loss condition

```{r}
t.test(subject_color_cse$CSE_loss, mu = 0, alternative = "greater")
```

### H2.3 - Congruency-sequence Effect for neutral condition

```{r}
t.test(subject_color_cse$CSE_neutral, mu = 0, alternative = "greater")
```

### Assumption check for H3

```{r}
assumption_h3 = with(subject_color_cse, CSE_loss - CSE_neutral)
shapiro.test(assumption_h3)
```

### H3 - CSE will be stronger in the loss condition compared to neutral (paired samples t-test)

```{r}
t.test(subject_color_cse$CSE_loss, subject_color_cse$CSE_neutral, paired = T, alternative = "greater")
```

### Assumption check for H4

```{r}
assumption_h4 = with(subject_color_cse, CSE_neutral - CSE_gain)
shapiro.test(assumption_h4)
```

### H4 - CSE will be weaker in the gain condition compared to neutral (paired samples t-test)

```{r}
t.test(subject_color_cse$CSE_gain, subject_color_cse$CSE_neutral, paired = T, alternative = "less")
```





