---
title: "109_cse_analysis_v2"
author: "Peter Czingraber"
date: "2026-06-19"
output: html_document
---

## loading packages

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(bayesplot)
library(tidybayes)
library(bayestestR)
library(posterior)
library(see)
library(bayesboot)
```

---

## Importing data

```{r}
raw_pilot_data = read_csv("raw_pilot_data.csv")
```

---

## Data cleaning

```{r}
demography_data = raw_pilot_data %>%
  filter(trial_type %in% c("survey-text", "survey-multi-choice")) %>%
  select(response, subj_code) %>%
  mutate(
    age = str_extract(response, '"age":"\\d+"') %>%
      str_extract("\\d+") %>%
      as.numeric(),
    gender = str_extract(response, '"gender":"[^"]+"') %>%
      str_remove_all('"gender":"') %>%
      str_remove('"')
  ) %>%
  select(-response) %>%
  group_by(subj_code) %>%
  summarize(
    age    = coalesce(age[1], age[2]),
    gender = coalesce(gender[1], gender[2])
  )

demography_data %>%
  summarise(mean_age = mean(age, na.rm = TRUE),
            n_female = sum(gender == "Nő", na.rm = TRUE))
```

---

## Flagging practice trials 

```{r}
raw_pilot_data = raw_pilot_data %>% 
  group_by(subj_code) %>% 
  mutate(block_marker = coalesce(str_detect(stimulus, regex("Blokk\\s*1\\s*kezdődik", ignore_case = TRUE)),FALSE),
         start_seen = cummax(trial_index == 0),
         marker_seen = cummax(block_marker),
         trial_exp = if_else(start_seen == 1 & marker_seen == 0, "practice", "experimental")) %>% 
  ungroup()
```

---

## Flagging previous trials

```{r}
raw_pilot_data = raw_pilot_data %>%
  mutate(prev_congruency = lag(congruency, n = 4),
         prev_correct = lag(correct, n = 4),
         correct = ifelse(correct == "TRUE", 1, 0),
         prev_correct = ifelse(prev_correct == "TRUE", 1, 0),
         first_trial = ifelse(lag(str_detect(stimulus, "Blokk \\d+ kezdődik"), 4) == TRUE, 1, 0),
         rt = as.numeric(rt))
```

---

## Flagging diagnostic trials

```{r}
raw_pilot_data = raw_pilot_data %>%
  filter(task == "probe", trial_exp == "experimental") %>%
  group_by(subj_code) %>%
  mutate(trial_role = if_else(monetary == 1, "inducer", "diagnostic")) %>%
  ungroup()
```

---

## Flagging monetary-monetary sequence

```{r}
# The trials were generated by a randomization python code with specific parameters. It is possible, that because of this randomness, it happened in a few cases that a monetary trial followed another monetary trial. This shouldn't be the case, therefore we exclude these pairs.

raw_pilot_data <- raw_pilot_data %>%
  group_by(subj_code) %>%
  mutate(double_monetary = if_else(
    monetary == 1 & (lag(monetary, default = 0) == 1 | lead(monetary, default = 0) == 1),
    1, 0
  )) %>%
  filter(double_monetary != 1) %>%
  ungroup()
```

---

## Filtering subjects with accuracy below 60%

```{r}
accuracy = raw_pilot_data %>%
  group_by(subj_code) %>%
  summarize(all_accuracy = mean(correct, na.rm = TRUE))

raw_pilot_data = raw_pilot_data %>%
  select(-any_of("all_accuracy")) %>%
  left_join(accuracy, by = "subj_code") %>%
  filter(all_accuracy > 0.60, first_trial != 1)
```

---

## Keeping correct trials only

```{r}
filtered_pilot_data = raw_pilot_data %>%
  filter(correct == 1, prev_correct == 1, rt >= 150 | is.na(rt))
```

---

## Omitting unnecessary variables

```{r}
processed_pilot_data = filtered_pilot_data %>%
  select(
    -experiment, -double_monetary, -stimulus, -response,
    -trial_type, -trial_index, -plugin_version, -time_elapsed,
    -success, -question_order, -correct_response, -block_marker,
    -marker_seen, -prev_correct, -start_seen, -trial_exp,
    -first_trial, -name, -money, -source_file
    )
```

---

## Flagging previous color and labels

```{r}
processed_pilot_data = processed_pilot_data %>%
  group_by(subj_code) %>%
  mutate(prev_color = lag(color)) %>%
  ungroup() %>%
  mutate(prev_color = case_when(
    prev_color == "red"                             ~ "loss",
    prev_color == "green"                           ~ "gain",
    prev_color %in% c("magenta", "blue", "yellow")  ~ "neutral"
  )) %>%
  filter(!is.na(prev_color), trial_role == "diagnostic")
```

```{r}
write.csv(processed_pilot_data, file = "processed_pilot_data.csv")
```

---

## Descriptive statistics and visualisations

```{r}
dstats = function(x, na.omit=FALSE){
  if (na.omit)
  x = x[!is.na(x)]
  min = min(x)
  max = max(x)
  m = mean(x)
  median = median(x)
  n = length(x)
  stds = sd(x)
  skew = 3*(m-median)/stds
  kurt = (length(x) * sum((x - mean(x))^4)) / (sum((x - mean(x))^2)^2) - 3

  
#Return all calculated values to vectors
  return(c(
    min = min,
    max = max,
    mean = m,
    median = median,
    sd = stds,
    skewness = skew,
    kurtosis = kurt
  ))
}
```

```{r}
processed_pilot_data %>% 
  summarise(n_above_1000 = sum(rt > 1000),
            percent_above_1000 = mean(rt > 1000)*100)
```


---

## Descriptive statistics for aggregated RT values

```{r}
mean_rt_df = processed_pilot_data %>% 
  group_by(subj_code) %>% 
  summarise(mean_rt_congruent = mean(rt[congruency == "congruent"]),
            mean_rt_incongruent = mean(rt[congruency == "incongruent"]))

myvars = c("mean_rt_congruent", "mean_rt_incongruent")
sapply(mean_rt_df[myvars], dstats)
```

---

## Density plot

```{r}
ggplot(processed_pilot_data, aes(rt)) +
  geom_density() +
  theme_minimal() +
  labs(title = "RT distribution (all diagnostic trials)")
```

---

## Violinplots for CE

```{r}
ce_rt = processed_pilot_data %>%
  group_by(subj_code, congruency) %>%
  summarise(mean_rt = mean(rt), .groups = "drop")

violin = ggplot(ce_rt, aes(congruency, mean_rt, fill = congruency)) +
  geom_violin(trim = TRUE, bounds = c(0, Inf), width = 0.5) +
  geom_boxplot(width = 0.1, colour = c("black", "white"), outlier.shape = NA) +
  labs(title = "Congruency Effect", x = "Congruency", y = "Mean reaction time (ms)") +
  papaja::theme_apa(base_family = "Times") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("congruent" = "#FAC775", "incongruent" = "#BA7517"))
#ggsave("ce_violin.png", plot = violin, width = 16, height = 9, units = "cm", dpi = 320)
print(violin)
```

---

## Participant level CSE

```{r}
# Appropriate data structure for visualizing the CSE
cse_df = processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency) %>% 
  summarise(mean_rt = mean(rt))

cse_descriptive_df = cse_df %>% 
  group_by(congruency, prev_congruency) %>% 
  summarise(N = n(),
            mean = mean(mean_rt),
            sd = sd(mean_rt),
            se = sd/sqrt(N))

wide_df = cse_df %>% 
  pivot_wider(names_from = c(prev_congruency, congruency),
              values_from = mean_rt)

subject_cse = wide_df %>% 
  mutate(CSE = (congruent_incongruent - congruent_congruent) - (incongruent_incongruent - incongruent_congruent))

ggplot(subject_cse, aes(CSE)) +
  geom_density(fill = "#011526", alpha = 0.25) +
  scale_x_continuous(breaks = seq(floor(min(subject_cse$CSE)),
                                  ceiling(max(subject_cse$CSE)),
                                  by = 5)) +
  theme_minimal()

ggplot(subject_cse, aes(CSE)) +
  geom_histogram(fill = "#011526", alpha = 0.25, binwidth = 10) +
  scale_x_continuous(breaks = seq(floor(min(subject_cse$CSE)),
                                  ceiling(max(subject_cse$CSE)),
                                  by = 5)) +
  theme_minimal()
```

---

## CSE plot

```{r}
cse_plot = ggplot(cse_descriptive_df, aes(prev_congruency, mean, colour = congruency, group = congruency)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1.2) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 1.2) +
                      labs(
                        title = "Congruency-sequence Effect",
                       x = "Previous Congruency",
                       y = "Mean reaction time (ms)",
                       colour = "Congruency") + 
  scale_colour_manual(values = c("#FAC775", "#BA7517")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right",
                        plot.title = element_text(hjust = 0.5, face = "bold"),
                        panel.grid = element_blank(),
                        axis.line = element_line(color = "grey", linewidth = 1))
#ggsave("overall_cse_plot.png", plot = cse_plot, width = 16, height = 9, units = "cm", dpi = 320)
print(cse_plot)
```

---

## CSE descriptive

```{r}
myvars = c("CSE")
sapply(subject_cse[myvars], dstats)
```

---

## Participant level CSE for the three different condition

```{r}
cse_color_df= processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency, prev_color) %>% 
  summarise(mean_rt = mean(rt))

wide_color_df = cse_color_df %>% 
  pivot_wider(names_from = c(prev_color, prev_congruency, congruency),
              values_from = mean_rt)

subject_color_cse = wide_color_df %>% 
  mutate(CSE_loss = (loss_congruent_incongruent - loss_congruent_congruent) - (loss_incongruent_incongruent - loss_incongruent_congruent),
         CSE_gain = (gain_congruent_incongruent - gain_congruent_congruent) - (gain_incongruent_incongruent - gain_incongruent_congruent),
         CSE_neutral = (neutral_congruent_incongruent - neutral_congruent_congruent) - (neutral_incongruent_incongruent - neutral_incongruent_congruent)) %>% 
  filter(!is.na(CSE_loss) & !is.na(CSE_gain) & !is.na(CSE_neutral))
```

---

## Descriptive stats for all CSE conditions

```{r}
myvars2 = c("CSE_loss", "CSE_gain", "CSE_neutral")
sapply(subject_color_cse[myvars2], dstats)
```

---

## Visualizing the CSE distributions

```{r setup, message=FALSE, warning=FALSE}
subject_color_cse %>%
  pivot_longer(cols = c(CSE_loss, CSE_gain, CSE_neutral),
               names_to = "condition",
               values_to = "CSE") %>%
  filter(!is.na(CSE)) %>%
  ggplot(aes(x = CSE, fill = condition)) +
  geom_histogram(alpha = 0.25, binwidth = 5, fill = "#011526") +
  facet_wrap(~condition) +
  scale_x_continuous(breaks = seq(-200, 200, by = 25))
  theme_minimal()
```

---

## Plotting CSE with emotional condition

```{r}
cse_color_summary_df = processed_pilot_data %>% 
  group_by(subj_code, congruency, prev_congruency, prev_color) %>% 
  summarise(mean_rt = mean(rt))

cse_color_plot_df = cse_color_summary_df %>% 
  group_by(congruency, prev_congruency, prev_color) %>% 
  summarise(N = n(),
            grand_mean_rt = mean(mean_rt),
            sd_rt = sd(mean_rt),
            se_rt = sd_rt/sqrt(N))

cse_color_plot_df %>% 
  ggplot()+
  aes(y = grand_mean_rt, x = prev_congruency, colour = prev_color, group = congruency)+
  geom_point(size = 3) +
  geom_path(linewidth = 1.2)+
  geom_errorbar(aes(ymin = grand_mean_rt - se_rt, ymax = grand_mean_rt + se_rt), width = 0.2, linewidth = 0.8) +
                  labs(title = "Congruency-sequence Effect",
                       x = "Previous Congruency",
                       y = "Mean reaction time (ms)",
                       colour = "Congruency") + 
                  scale_colour_manual(values = c("#03AC13", "#B90E0A", "#616569")) +
                  theme_minimal(base_size = 14) +
                  theme(legend.position = "right",
                        plot.title = element_text(hjust = 0.5, face = "bold"),
                        panel.grid = element_blank(),
                        axis.line = element_line(color = "grey", linewidth = 1)) +
  facet_grid(cols = vars (prev_color))
#ggsave("cse_color_plot.png", plot = last_plot(), width = 16, height = 9, units = "cm", dpi = 320)
```

---

## Hierarchical ExGaussian Model

### Contrast coding

```{r}
# All three predictors are sum-coded so that:
#   - The intercept = grand mean RT across all conditions (in raw ms)
#   - Every coefficient = a contrast around that grand mean, also in raw ms
#   - The three-way interaction = the difference in CSE between monetary conditions, ídirectly readable in milliseconds
#
# Coding of the 3-level monetary condition
#
# We use two centered (sum-to-zero) contrasts:
#
#   c_loss: loss = +2/3, gain = -1/3, neutral = -1/3
#   c_gain: loss = -1/3, gain = +2/3, neutral = -1/3
#
# IMPORTANT:
#
# These contrasts are centered (their values sum to zero), so the model intercept remains the grand mean across all three monetary conditions. However, c_loss and c_gain are NOT orthogonal to each other. Their dot product is:
#
# = (2/3)(-1/3) + (-1/3)(2/3) + (-1/3)(-1/3)
# = -2/9 -2/9 +1/9
# = -1/3
#
# Because both contrasts are entered into the model simultaneously, every coefficient is interpreted as a PARTIAL regression coefficient. That is after statistically controlling for the other monetary contrast.
#
# Consequently:
#
#   b_curr_cong:prev_cong:c_loss
#
# DOES NOT compare the loss condition against the average of gain and neutral, despite the appearance of the coding. Instead, after controlling for c_gain, the gain condition contributes zero unique information to the c_loss coefficient, making the comparison effectively:
#
#   Loss CSE  vs.  Neutral CSE
#
# Same reasoning applies for the gain condition
#
# Thus the three-way interaction coefficients are directly interpretable as raw millisecond differences relative to the neutral condition, with no additional scaling factor required.
#
# curr_cong:    congruent = -0.5, incongruent = +0.5
# prev_cong:    congruent = +0.5, incongruent = -0.5
# prev_color:   c_loss: loss = +2/3, gain = -1/3, neutral = -1/3
#               c_gain: loss = -1/3, gain = +2/3, neutral = -1/3
#
# With this coding:
#   b_curr_cong                  = CE (incongruent - congruent)                               [H1]
#   b_curr_cong:prev_cong        = the CSE interaction term                                   [H2]
#   b_curr_cong:prev_cong:c_loss = CSE modulation by loss vs neutral (controlling for gain)   [H3]
#   b_curr_cong:prev_cong:c_gain = CSE modulation by gain vs neutral (controlling for loss)   [H4]

model_data = processed_pilot_data %>%
  mutate(
    curr_cong = if_else(congruency      == "incongruent", 0.5, -0.5),
    prev_cong = if_else(prev_congruency == "incongruent", -0.5, 0.5),  

    # Sum coding for the 3-level monetary condition
    c_loss = case_when(
      prev_color == "loss"    ~  2/3,
      prev_color == "gain"    ~ -1/3,
      prev_color == "neutral" ~ -1/3
    ),
    c_gain = case_when(
      prev_color == "loss"    ~ -1/3,
      prev_color == "gain"    ~  2/3,
      prev_color == "neutral" ~ -1/3
    ),

    subj_code = as.factor(subj_code)
  )
```

### Priors

```{r}
# Priors are on the raw millisecond scale (using identity link).
#
# Intercept ~ Normal(500, 100): centered on a plausible grand-mean RT, with generous uncertainty.
#
# b (fixed effects) ~ Normal(0, 50): allows for effects up to ~100ms, which is well above any plausible CSE modulation. Weakly informative prior for all regression coefficients
#
# sd (random effect SDs) ~ Normal(0, 50): same reasoning; individual differences even in the CSE can drastically vary.
#
# sigma ~ Normal(0, 100): the Gaussian spread of the ExGaussian
#
# beta ~ Normal(0, 100): the exponential tail of the ExGaussian
# 
# In our previous experiments utilizing similar prime probe designs, the rt distributions suggest that sigma is in the 60-80 ms range. Also the rt distributions show a gradual tail extending to ~1000 ms in this prime-probe design, suggesting that beta is in the 80-120 ms range. Therefore setting both of these parameters to SD = 100 ms could reflect this, while remaining weakly informative. Both of these parameters are constrained to be positive by brms automatically.

exg_priors = c(
  prior(normal(500, 100), class = Intercept),
  prior(normal(0, 50),   class = b, coef = "curr_cong"),
  prior(normal(0, 25),   class = b, coef = "curr_cong:prev_cong"),
  prior(normal(0, 12.5), class = b, coef = "curr_cong:prev_cong:c_loss"),
  prior(normal(0, 12.5), class = b, coef = "curr_cong:prev_cong:c_gain"),
  prior(normal(0, 50),   class = b),  # fallback for every other main effect
  prior(normal(0, 50),    class = sd),
  prior(normal(0, 100),   class = sigma),
  prior(normal(0, 100),   class = beta)
)
```

### Model formula

```{r}
# The full three-way interaction with all lower-order terms.
# This single model covers H1 through H4.
#
# Random effects: by-participant intercepts and slopes for curr_cong. This captures individual variation in CE size.
# A fully crossed random structure (including monetary slopes) is theoretically ideal but typically does not converge at realistic N; Perhaps we should estimate individual differences in CSE. Simplify further if convergence issues arise

exg_formula = bf(rt ~ curr_cong * prev_cong * c_loss + curr_cong * prev_cong * c_gain +
       (curr_cong | subj_code), family = exgaussian (link = "identity"))
```

### Fitting the model

```{r}
fit_exg = brm(
  formula = exg_formula,
  data    = model_data,
  prior   = exg_priors,
  chains  = 4,
  cores   = 4,
  iter    = 6000,
  warmup  = 3000,
  control = list(adapt_delta = 0.95), # up from the default 0.8: prior fit showed relatively low ESS/high Rhat on the
                                      # intercept only; smaller steps improve mixing here.
  seed    = 109,
  file    = "fit_exg_model"  # caches the fit; delete the file to refit
)
```

---

## Model diagnostics

```{r}
# All Rhat should be < 1.01; bulk and tail ESS should be > 400 per chain
summary(fit_exg)
```

```{r}
mcmc_trace(fit_exg,
           pars = c("b_Intercept",
                    "b_curr_cong",
                    "b_curr_cong:prev_cong",
                    "b_curr_cong:prev_cong:c_loss",
                    "b_curr_cong:prev_cong:c_gain"))
```

```{r}
# Posterior predictive check: does the model reproduce the RT distribution shape?
pp_check(fit_exg, ndraws = 100, type = "dens_overlay")
```

---

## Posterior summaries and HDIs for H1–H4

```{r}
draws = as_draws_df(fit_exg)

summarise_posterior = function(x, direction = "positive") {
  h = hdi(x, ci = 0.95)
  p_direction = if (direction == "positive") mean(x > 0) else mean(x < 0)
  tibble(
    mean      = mean(x),
    hdi_lower = h$CI_low,
    hdi_upper = h$CI_high,
    hdi_width = h$CI_high - h$CI_low,
    p_direction = p_direction
  )
}
```

---

## Pre-specified precision and inference criteria

```{r}
# H1 (CE):               30 ms — the CE is large and stable, wide target sufficient
# H2 (CSE):              20 ms — core effect of interest, tighter target
# H3 (loss modulation):  15 ms — critical contrast, high precision needed
# H4 (gain modulation):  15 ms — critical contrast, high precision needed
 
target_widths = list(H1 = 30, H2 = 20, H3 = 15, H4 = 15) 

# ROPE: inference criterion
# Effects smaller than ±5ms are considered practically negligible.
# If the 95% HDI falls entirely within [-5, 5] ms, the effect is absent for practical purposes.
 
rope_lower = -5
rope_upper =  5
```

## ── H1: Congruency Effect ──────────────────────────────────────

```{r}
# b_curr_cong: (incongruent - congruent) 

h1_draws   = draws$b_curr_cong 
h1_summary = summarise_posterior(h1_draws, direction = "positive")

cat("H1 — Congruency Effect (ms)\n")
print(h1_summary)
```

---

## ── H2: Congruency Sequence Effect ────────────────────────────

```{r}
# b_curr_cong:prev_cong: the interaction of the two ±0.5 predictors.
# This is the CSE — the difference in CE as a function of previous congruency.

h2_draws   = draws$`b_curr_cong:prev_cong`
h2_summary = summarise_posterior(h2_draws, direction = "positive")

cat("H2 — Congruency Sequence Effect (ms)\n")
print(h2_summary)
```

---

## ── H3: Loss enhances CSE ────────────────────────────────────

```{r}
# b_curr_cong:prev_cong:c_loss: how much the CSE differs in the loss condition relative to the neutral condition.
# A positive value = larger CSE after loss inducers.

h3_draws = draws$`b_curr_cong:prev_cong:c_loss`
h3_summary = summarise_posterior(h3_draws, direction = "positive")

cat("H3 — Loss enhances CSE vs neutral (ms)\n")
print(h3_summary)
```

---

## ── H4: Gain reduces CSE ────────────────────────────────────

```{r}
# b_curr_cong:prev_cong:c_gain: how much the CSE differs in the gain condition relative to the neutral condition. 
# A negative value = smaller CSE after gain inducers.

h4_draws = draws$`b_curr_cong:prev_cong:c_gain`
h4_summary = summarise_posterior(h4_draws, direction = "negative")

cat("H4 — Gain reduces CSE vs neutral (ms)\n")
print(h4_summary)
```

---

## Precision and ROPE results table

```{r}
# Two separate criteria:
#   1. Precision: is the HDI width below the target? (stopping rule)
#   2. Inference: where does the HDI fall relative to the ROPE? (inference rule)
 
eq_h1 = equivalence_test(h1_draws, range = c(rope_lower, rope_upper), ci = 0.95)
eq_h2 = equivalence_test(h2_draws, range = c(rope_lower, rope_upper), ci = 0.95)
eq_h3 = equivalence_test(h3_draws, range = c(rope_lower, rope_upper), ci = 0.95)
eq_h4 = equivalence_test(h4_draws, range = c(rope_lower, rope_upper), ci = 0.95)

results = tibble(
  hypothesis    = c("H1 (CE)", "H2 (CSE)", "H3 (loss)", "H4 (gain)"),
  hdi_lower     = c(h1_summary$hdi_lower, h2_summary$hdi_lower,
                     h3_summary$hdi_lower, h4_summary$hdi_lower),
  hdi_upper     = c(h1_summary$hdi_upper, h2_summary$hdi_upper,
                     h3_summary$hdi_upper, h4_summary$hdi_upper),
  hdi_width     = c(h1_summary$hdi_width, h2_summary$hdi_width,
                     h3_summary$hdi_width, h4_summary$hdi_width),
  target        = c(target_widths$H1, target_widths$H2,
                     target_widths$H3, target_widths$H4),
  precision_met = hdi_width < target,
  decision      = c(eq_h1$ROPE_Equivalence, eq_h2$ROPE_Equivalence,
                     eq_h3$ROPE_Equivalence, eq_h4$ROPE_Equivalence)
)

print(results)
```

---

## Plots

```{r}
# Each plot shows the posterior distribution with the 95% HDI and ROPE marked.
# They use the same equivalence_test() object as the results table, so the visual and the table will match.

# H1
plotPost(h1_draws, credMass = 0.95, compVal = 0, ROPE = c(rope_lower, rope_upper), showCurve = TRUE, xlab = "CE (ms)", main = "H1 — Congruency Effect")

# H2
plotPost(h2_draws, credMass = 0.95, compVal = 0, ROPE = c(rope_lower, rope_upper), showCurve = TRUE, xlab = "CSE (ms)", main = "H2 — Congruency Sequence Effect")

# H3
plotPost(h3_draws, credMass = 0.95, compVal = 0, ROPE = c(rope_lower, rope_upper), showCurve = TRUE, xlab = "Loss modulation (ms)", main = "H3 — Loss enhances CSE")

# H4
plotPost(h4_draws, credMass = 0.95, compVal = 0, ROPE = c(rope_lower, rope_upper), showCurve = TRUE, xlab = "Gain modulation (ms)", main = "H4 — Gain reduces CSE")
```

---

## Manipulation checks (H5–H8) - NOT YET WORKED OUT
### Normal-likelihood Bayesian models on VAS scores (unchanged approach, updated to brms)

```{r}
# VAS scores are bounded continuous data on [-50, +50].
# A Gaussian likelihood is a reasonable approximation here.
# Structure mirrors the RT model: fixed effect of condition,
# by-participant random intercepts.
#
# Uncomment and adapt when VAS data are available.

# vas_priors <- c(
#   prior(normal(0, 25),  class = Intercept),
#   prior(normal(0, 25),  class = b),
#   prior(normal(0, 25),  class = sd),
#   prior(normal(0, 25),  class = sigma)
# )
#
# fit_valence <- brm(
#   valence ~ condition + (1 | subj_code),
#   data   = valence_VAS_long,
#   family = gaussian(),
#   prior  = vas_priors,
#   chains = 4, cores = 4, iter = 4000, warmup = 2000, seed = 109,
#   file   = "fit_valence"
# )
#
# fit_arousal <- brm(
#   arousal ~ condition + (1 | subj_code),
#   data   = arousal_VAS_long,
#   family = gaussian(),
#   prior  = vas_priors,
#   chains = 4, cores = 4, iter = 4000, warmup = 2000, seed = 109,
#   file   = "fit_arousal"
# )
```



